Visualizing probability distributions across bivariate cyclic temporal granularities

### 2020-04-02 main.Rmd
rank harmonies and ppt diagram update

### Unevenly distributed events

@@

@@

R package **gravitas** [@R-gravitas] is used to facilitate the systematic exploration here. While trying to explore the energy behavior of these customers systematically across cyclic time granularities, the first thing we should have at our disposal is to know which all cyclic time granularities we can look at exhaustively. Let us consider conventional time deconstructions for a Gregorian calendar (second, minute, half-hour, hour, day, week, month, year). Since the interval of this tsibble is 30 minutes, the temporal granularities may range from half-hour to year. Considering $6$ linear granularities half-hour, hour, day, week, month, year, there could be $(6*5/2) = 15$ circular or quasi-circular granularities that could be formed. If these options are considered too many, the most coarse temporal unit can be set to a “month” and some intermediate temporal units which might not be pertinent to the analysis like hhour could be dropped. Then we will be left with six cyclic granularites namely "hour_day", "hour_week", "hour_month", "day_week", "day_month" and "week_month".

## Computation

Now that we have a list of cyclic granularities to look at, we should be able to compute them from the data using Sections \ref{sec:cyclic-calendar}, \ref{sec:circular-gran-def} (for hour_day, hour_week and day_week) and \ref{sec:quasi-circular-gran-def} or (for hour_month, day_month, week_month).

@@

```{r harmony-tab, echo=FALSE}
smart_harmony <- smart_meter10 %>% 
  harmony(ugran = "month",
          filter_out = c("hhour", "fortnight"))%>%
  rename(`facet variable` = facet_variable,
         `x-axis variable` = x_variable,
         `Levels of facet variable` = facet_levels,
         `Levels of x-axis variable` = x_levels)

  knitr::kable(smart_harmony,
               format = "latex",
               booktabs = TRUE,
                caption = "Harmonies with two cyclic granularity one placed on facet and the other on x-axis along with their number of categories/levels. Out of 30 possible combinations of cyclic granularities, only these 13 are harmony pairs.") %>% 
    #row_spec(0, bold = TRUE) %>% 
    kable_styling()
```

>>>>

@@

## Ranking harmonies based on total variability

Even after screening harmonies from the list of all possible combinations of cyclic time granularities that could be plotted together, the options could be too many and it might be useful to rank these harmonies. Ranking could be based on how well a harmony pair displays the maximum variation in distributions of the measured variable within and across each levels of the cyclic granularities. For each facet level, the most important display is the one in which the difference in the distribution of the measured variable between the x-axis categories is maximum. And the harmony pair which has more variation between x-axis levels for all facet levels on an average is likely to be a good measure of how important a harmony pair is displaying the insights of the data. Distances between probability distributions of the measured variable across x-axis levels are computed using Jenson-Shannon distances which is a symmetric divergence obtained from Kullback-Leiber divergence measure. Rpobability distributions are represented through sample percentiles instead of kernel density estimate so that there is minimal dependency on selecting kernel or bandwidth.

@@

R package **gravitas** [@R-gravitas] is used to facilitate the systematic exploration here. While trying to explore the energy behavior of these customers systematically across cyclic time granularities, the first thing we should have at our disposal is to know which all cyclic time granularities we can look at exhaustively. Let us consider conventional time deconstructions for a Gregorian calendar (second, minute, half-hour, hour, day, week, month, year). Since the interval of this tsibble is 30 minutes, the temporal granularities may range from half-hour to year. Considering $6$ linear granularities half-hour, hour, day, week, month, year, there could be $(6*5/2) = 15$ circular or quasi-circular granularities that could be formed. If these options are considered too many, the most coarse temporal unit can be set to a “month” and some intermediate temporal units which might not be pertinent to the analysis like hhour could be dropped. Then we will be left with six cyclic granularites namely "hour_day", "hour_week", "hour_month", "day_week", "day_month" and "week_month". Now that we have a list of cyclic granularities to look at, we should be able to compute them from the data using Sections \ref{sec:cyclic-calendar}, \ref{sec:circular-gran-def} (for hour_day, hour_week and day_week) and \ref{sec:quasi-circular-gran-def} (for hour_month, day_month, week_month).

@@

```{r harmony-tab, echo=FALSE}
harmonies <- smart_meter10 %>% 
  harmony(ugran = "month",
          filter_out = c("hhour", "fortnight"))

smart_harmony <- smart_meter10 %>%
  filter(customer_id %in% c(10017936)) %>% 
  rank_harmony(harmony_tbl = harmonies,response =  "general_supply_kwh")%>%
  rename(`facet variable` = facet_variable,
         `x-axis variable` = x_variable,
         `facet levels` = facet_levels,
         `x-axis levels` = x_levels,
         `average maximum variation` = mean_max_variation)

  knitr::kable(smart_harmony,
               format = "latex",
               booktabs = TRUE,
                caption = "Harmonies with two cyclic granularity one placed on facet and the other on x-axis along with their number of categories/levels. Out of 30 possible combinations of cyclic granularities, only these 13 are harmony pairs. The harmony pairs are ranked basis the average maximum variation which captures maximum pairwise distance between x-axis levels for each facets and then average them over facets.") %>% 
    #row_spec(0, bold = TRUE) %>% 
    kable_styling()
```
